Active Learning for SAT Solver Benchmarking

### 2024-05-24 paper_jar/paper.tex
section 4

Given all the previously presented instantiations for Algorithm~\ref{algALBenchmark}, this section outlines our experimental design, including our evaluation framework, used data sets, hyper-parameter choices, and implementation details.

@@

  $O \leftarrow \emptyset$
  \BlankLine

@@

  \BlankLine
    $O \leftarrow O \cup \bigl\{ ( O_{\operatorname{acc}},\, O_{\operatorname{rt}} ) \bigr\}$
  }

  \BlankLine
  $\bigl( \bar{O}_{\operatorname{acc}}, \bar{O}_{\operatorname{rt}} \bigr) \leftarrow \operatorname{average}(O)$ \;

  \BlankLine
  \Return $\bigl( \bar{O}_{\operatorname{acc}}, \bar{O}_{\operatorname{rt}} \bigr)$
\end{algorithm}

As stated in the Introduction, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
As described in Section~\ref{sec:main:model}, a prediction model $\mathcal{M}$ provides us with an estimated scoring~$s_{\hat a}$ for the new solver~$\hat{a}$.

To evaluate a concrete instantiation of Algorithm~\ref{algALBenchmark}, i.e., a concrete choice for all the sub-routines, we perform cross-validation on our set of solvers.
Algorithm~\ref{alg:eval} shows this.
That means each solver plays the role of the new solver~$\hat{a}$ once (Line~2).
Note that the \emph{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage (Line~3).
After running our active-learning framework for solver~$\hat{a}$ (Line~4), we compute the value of both our optimization goals, i.e., ranking accuracy and runtime.
We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly regarding the ground-truth scoring~$\operatorname{par_2}$ (Lines~5-8).
The \emph{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset (Lines~9-13).
Finally, we compute averages of the output metrics in Line~15 after we have collected all cross-validation results in Line~14.
Overall, we want to find an approach that maximizes
%
\begin{equation}
  O_\delta := \delta O_{\operatorname{acc}} + \left(1 - \delta\right) \left(1 - O_{\operatorname{rt}}\right) \enspace \textrm{,}
  \label{eq:opt}
\end{equation} 
%
whereby $\delta \in \left[0, 1\right]$ allows for linear weighting between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on an $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.

\subsection{Data}
\label{sec:exdesign:data}

In our experiments, we work with the dataset of the SAT~Competition~2022 Anniversary Track~\cite{sat2022}.
The dataset consists of 5355 instances with respective runtime data of 28 sequential SAT solvers.
We also use a database of 56 instance features\footnote{\url{https://benchmark-database.de/getdatabase/base_db}} from the Global Benchmark Database~(GBD) by Iser~et~al.~\cite{IserS18}.
They comprise instance size features and node distribution statistics for several graph representations of SAT instances, among others, and are primarily inspired by the SATzilla~2012 features described in~\cite{features}.
All features are numeric and free of missing values.
We drop 10 out of 56 features because of zero variance.
Overall, prediction models have access to 46 instance features and 27 runtime features, i.e., excluding the current new solver~$\hat{a}$.

Additionally, we retrieve instance-family information\footnote{\url{https://benchmark-database.de/getdatabase/meta_db}} to evaluate the composition of our sampled benchmarks.
Instance families comprise instances from the same application domain, e.g., planning, cryptography, etc., and are a valuable tool for analyzing solver performance.

For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instances' family.
All instance families that are too \emph{small}, i.e., \SI{10}{\%} of them corresponds to less than one instance, are put into one meta-family for stratification.
This \emph{tuning dataset} allows for a more extensive exploration of the hyper-parameter space.

\subsection{Hyper-parameters}
\label{sec:exdesign:hyper}

Given Algorithm~\ref{algALBenchmark}, there are several possible instantiations for the three sub-routines, i.e., \emph{ranking}, \emph{selection}, and \emph{stopping}.
We describe these experimental configurations in the following.

>>>>

In light of the preceding instantiations of Algorithm~\ref{algALBenchmark}, this section outlines the experimental design employed in the presented study.
This includes an overview of the evaluation framework, the data sets employed, the hyper-parameter selections, and the implementation details.

@@

  $O \leftarrow \emptyset$
  % \BlankLine

@@

  % \BlankLine
    $O \leftarrow O \cup \bigl\{ ( O_{\operatorname{acc}},\, O_{\operatorname{rt}} ) \bigr\}$
  }

  % \BlankLine
  $\bigl( \bar{O}_{\operatorname{acc}}, \bar{O}_{\operatorname{rt}} \bigr) \leftarrow \operatorname{average}(O)$ \;

  % \BlankLine
  \Return $\bigl( \bar{O}_{\operatorname{acc}}, \bar{O}_{\operatorname{rt}} \bigr)$
\end{algorithm}

As stated in the Introduction, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
As described in Section~\ref{sec:main:model}, a prediction model provides us with an estimated scoring value for the new solver~$\hat{a}$.
%
To evaluate the efficacy of a concrete instantiation of Algorithm~\ref{algALBenchmark}, which entails a specific choice for all subroutines, we perform cross-validation on our set of solvers.

This is demonstrated in Algorithm~\ref{alg:eval}.
Each solver assumes the role of the new solver~$\hat{a}$ once (Line~2) and is consequently excluded from the set of solvers $\mathcal{A}$ (Line~3).
After running our active-learning framework for solver~$\hat{a}$ (Line~4), we compute the value of both our optimization goals, i.e., ranking accuracy and runtime.
%
The \emph{ranking accuracy}, expressed in the interval $[0,1]$ (higher is better), is defined as the fraction of correctly ranked pairs of elements, namely, $\left(\hat{a}, a\right)$, where $a$ belongs to the set of all solvers $\mathcal{A}$ ranked by the ground-truth scoring function $\operatorname{par_2}$ (Lines~5-8).
%
The \emph{fraction of runtime} required by the algorithm to reach its conclusion is represented by $O_{\operatorname{rt}}$, which lies within the interval $[0, 1]$ (lower is better).
This metric compares the accumulated runtimes over the sampled instances to the accumulated runtimes over all instances in the dataset (Lines~9-13).
Once all cross-validation results have been collected (Line~14), the output metrics are averaged (Line~15).

The procedure is designed to identify an approach that maximizes Equation~\eqref{eq:opt}, in which $\delta \in \left[0, 1\right]$ is used for linear weighting, with weighting factor $\delta \in \left[0, 1\right]$, between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
The plotting of the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on an $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.
%
\begin{equation}
  O_\delta := \delta O_{\operatorname{acc}} + \left(1 - \delta\right) \left(1 - O_{\operatorname{rt}}\right) \enspace \textrm{,}
  \label{eq:opt}
\end{equation} 
%
%
\subsection{Data}
\label{sec:exdesign:data}

In our experiments we work with the dataset of the SAT~Competition~2022 Anniversary Track~\cite{sat2022}.
The dataset consists of 5355 instances with corresponding runtime data of 28 sequential SAT solvers.
We also use a database of 56 instance features\footnote{\url{https://benchmark-database.de/getdatabase/base_db}} from the Global Benchmark Database~(GBD) by Iser~et~al.~\cite{IserS18}.
They include instance size features and node distribution statistics for several graph representations of SAT instances, and are primarily inspired by the SATzilla~2012 features described in~\cite{features}.
All features are numeric and have no missing values.
We drop 10 of the 56 features due to zero variance.
In total, the prediction models have access to 46 instance features and 27 out of 28 runtime features, excluding the respective new solver~$\hat{a}$.

In addition, we retrieve instance family information\footnote{\url{https://benchmark-database.de/getdatabase/meta_db}} to evaluate the composition of our sampled benchmarks.
Instance families consist of instances from the same application domain, such as planning, cryptography, etc., and are a valuable tool for analyzing solver performance.
%
For hyperparameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance family.
All instance families that are too \emph{small}, i.e., \SI{10}{\%} of them correspond to less than one instance, are put into a meta-family for stratification.
This \emph{tuning dataset} allows a more extensive exploration of the hyperparameter space.

\subsection{Hyperparameters}
\label{sec:exdesign:hyper}

Given the Algorithm~\ref{algALBenchmark}, there are several possible instantiations for the three subroutines \emph{ranking}, \emph{selection}, and \emph{stopping}.
These experimental configurations are described in the following.
